---
title: "P8150 homework 2"
author: Yu He
Date: 10/6/2021
output: github_document
---

## homework 2 Problem 1
initla package

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

import and manipulate Mr. Trash Wheel data

```{r}
Mtw_df <- read_excel("hw2_data/Trash_Wheel_Collection.xlsx", sheet = 1, range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  filter(!(month %in% c("January Total", "February Total", "March Total", "April Total", "May Total", "June Total", "July Total", "August Total", "September Total", "October Total", "November  Total", "December Total", "Grand Total"))) %>% 
  mutate(sports_balls = round(sports_balls))
  
```

import and manipulate Mr. Trash Wheel precipitation data for 2018 and 2019

```{r}
Twp2019_df <- read_excel("hw2_data/Trash_Wheel_Collection.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)
  

Twp2018_df <- read_excel("hw2_data/Trash_Wheel_Collection.xlsx", sheet = "2018 Precipitation",  range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

Twp_df <- bind_rows(Twp2019_df, Twp2018_df) %>% 
  mutate(month = month.name[as.numeric(month)])
  
```

The Trash_Wheel_Collection data set include information collected from May, 2014 to January, 2021. It provides information about variables includes `r names(Mtw_df)`.

The average weight of trash collected each month is `r mean(pull(Mtw_df, weight_tons))`, and the average volume of trash collected each month is `r mean(pull(Mtw_df, volume_cubic_yards))`. In addition, the cigarette butts has the highest amount. which is `r sum(pull(Mtw_df, cigarette_butts))`.

The max precipitation happened in September, 2019, which is `r max(pull(Twp_df, total))`. The total precipitation in 2018 is `r sum(pull(filter(Twp_df, year == 2018), total))`. And the median number of sports balls in a dumpster in 2019 is`r median(pull(filter(Mtw_df, year == 2019), sports_balls))`.

## homework 2 Problem 2

Part 1 clean the data in pols-month. 
```{r}
pols_month <- read_csv("hw2_data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = prez_gop - prez_dem) %>% 
  mutate(president = recode(president, "1" = "gop", "2" = "gop", "-1" = "dem")) %>% 
  select(-prez_dem) %>% 
  select(-prez_gop) %>% 
  select(-day)
```


Part 2 clean the data in snp.csv

```{r}
snp_df = read_csv("hw2_data/snp.csv") %>% 
  janitor::clean_names() %>% 

  lubridate::mdy(date)
  
  separate(date, c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)])
  
```

Part 3 tidy the unemployment data with pols and snp data

```{r}
unemployment_df = read_csv("hw2_data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(-year, names_to = "month", values_to = "percentage of unemployment" ) %>%   
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"))
 
```

Part 4 merge all three tables

```{r}
fte_df = left_join(snp_df, pols_month) %>% 
  left_join(unemployment_df)
```

Add description:




## homework 2 problem 3
Part 1 load and tidy the data

```{r}
pb_names = read_csv("hw2_data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, "WHITE NON HISP" = "WHITE NON HISPANIC", "BLACK NON HISP" = "BLACK NON HISPANIC", "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(childs_first_name = tolower(childs_first_name)) %>% 
  distinct() 

```

Part 2(a) produce table for Baby's name Oliva

```{r}
filter(pb_names, childs_first_name == "olivia") %>%
  pivot_wider(names_from = year_of_birth, values_from = childs_first_name) %>% 
  knitr::kable()
```

Part 2(b) produce table for male children

```{r}
filter(pb_names, gender == "MALE") %>% 
  pivot_wider(names_from = year_of_birth, values_from = count) %>% 
  knitr::kable()
```

Part 3 produce a scatter plot for male
```{r}
filter(pb_names, year_of_birth == "2016") %>% 
  filter(ethnicity == "WHITE NON HISPANIC") %>% 
  filter(gender == "MALE") %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()
```


























